---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# scfaReporter

<!-- badges: start -->
<!-- badges: end -->

The goal of scfaReporter is to ...

## Installation

Install the latest internal build directly from GitHub:

```r
# install.packages("pak")
pak::pak("MicrobiomeInsights/scfaReporter")
# or: remotes::install_github("MicrobiomeInsights/scfaReporter")
```

## Quick start: run the pipeline on demo data

Two tiny demo datasets ship with the package so you can exercise the pipeline
without external files.

```{r example}
library(scfaReporter)

# Demo dataset 1: simple two-group comparison
demo1 <- run_scfa_pipeline(
  scfa_data = demo_scfa_dataset1,
  manifest = demo_manifest_dataset1,
  analytes = c("Acetic.Acid", "Propionic.Acid", "Butyric.Acid"),
  stats = "anova",
  aov_formula = value ~ Group
)

# Demo dataset 2: longitudinal design
demo2 <- run_scfa_pipeline(
  scfa_data = demo_scfa_dataset2,
  manifest = demo_manifest_dataset2,
  analytes = c("Acetic.Acid", "Propionic.Acid", "Butyric.Acid"),
  stats = "anova",
  aov_formula = value ~ Group * Time
)

head(demo1$summary)
demo2$plots$Butyric.Acid
```

## Generate the SCFA report

Render the bundled Quarto template directly from R:

```{r render-demo, eval=FALSE}
library(scfaReporter)

# Demo data (no Chromeleon files required)
render_scfa_report(
  output_file = "demo1.html",
  project_id = "Demo dataset 1",
  scfa_data = demo_scfa_dataset1,
  manifest = demo_manifest_dataset1,
  analytes = c("Acetic.Acid", "Propionic.Acid", "Butyric.Acid"),
  stats = "anova",
  aov_formula = value ~ Group
)

# Real Chromeleon exports + manifest
render_scfa_report(
  output_file = "client_project.html",
  project_id = "P-0123",
  scfa_dir = "data/chromatograms",
  manifest_path = "data/manifest.xlsx",
  normalization_multiplier = 7,
  normalization_basis = "mmol_per_L_plasma",
  stats = "lmer",
  lmer_formula = value ~ Group * Time + (1 | Subject_ID)
)
```

Prefer to edit the template locally? Draft it once and render manually:

1. Draft:
   ```r
   rmarkdown::draft(
     file = "reports/my_project.qmd",
     template = "scfa-report",
     package = "scfaReporter",
     create_dir = TRUE
   )
   ```
2. Update the YAML `params` in `reports/my_project/my_project.qmd` (set either
   `scfa_dir`/`manifest_path` or `scfa_data`/`manifest_override`).
3. Render: `quarto::quarto_render("reports/my_project/my_project.qmd")`.
   The HTML report (and optional `scfa_concentration_normalized.csv`) appears
   beside the `.qmd`.
