---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# scfaReporter

`scfaReporter` is an R package for processing short-chain fatty acid (SCFA)
chromatography exports, joining them to sample metadata, running common
statistical summaries, and rendering analysis-ready HTML reports.

The package is designed for workflows that start with Thermo Chromeleon exports
and a sample manifest, then produce normalized SCFA concentration tables,
summary statistics, plots, model results, and a Quarto report.

## Installation

Install the package from GitHub with `pak` or `remotes`:

```r
# install.packages("pak")
pak::pak("bcpd/scfa-reporter")

# or:
# install.packages("remotes")
remotes::install_github("bcpd/scfa-reporter")
```

## What It Does

- Imports SCFA chromatogram exports from a directory of `.xls` files.
- Reads and cleans sample manifest workbooks.
- Merges SCFA concentrations with sample metadata.
- Normalizes analyte concentrations with a configurable multiplier and basis.
- Runs ANOVA, linear model, or mixed-effects model workflows.
- Creates analyte-level summaries and plots.
- Renders a bundled Quarto HTML report.

## Quick Start

Two demo datasets are included so you can run the pipeline without external
Chromeleon files.

```{r example}
library(scfaReporter)

demo1 <- run_scfa_pipeline(
  scfa_data = demo_scfa_dataset1,
  manifest = demo_manifest_dataset1,
  analytes = c("acetic_acid", "propionic_acid", "butyric_acid"),
  stats = "anova",
  aov_formula = value ~ group
)

head(demo1$summary)
demo1$plots$butyric_acid
```

## Run the Pipeline on Project Files

For a real project, provide a directory of Chromeleon exports and a sample
manifest:

```r
library(scfaReporter)

results <- run_scfa_pipeline(
  scfa_dir = "data/chromatograms",
  manifest_path = "data/manifest.xlsx",
  normalization_multiplier = 7,
  normalization_basis = "mmol_per_L_plasma",
  stats = "lmer",
  lmer_formula = value ~ group * time + (1 | subject_id)
)

results$summary
results$plots$acetic_acid
```

If the manifest uses a non-standard sheet name or has extra header rows, read it
first and pass the cleaned object to the pipeline:

```r
manifest <- read_scfa_manifest(
  path = "data/manifest.xlsx",
  sheet = "Samples",
  skip = 20
)

results <- run_scfa_pipeline(
  scfa_dir = "data/chromatograms",
  manifest = manifest,
  stats = "anova",
  aov_formula = value ~ group
)
```

## Render an HTML Report

Use `render_scfa_report()` to run the workflow and render the bundled Quarto
template in one step.

```r
render_scfa_report(
  output_file = "scfa_report.html",
  project_id = "P-0123",
  scfa_dir = "data/chromatograms",
  manifest_path = "data/manifest.xlsx",
  normalization_multiplier = 7,
  normalization_basis = "mmol_per_L_plasma",
  stats = "lmer",
  lmer_formula = value ~ group * time + (1 | subject_id)
)
```

The report can also be rendered from in-memory demo data:

```r
render_scfa_report(
  output_file = "demo_report.html",
  project_id = "Demo dataset",
  scfa_data = demo_scfa_dataset1,
  manifest = demo_manifest_dataset1,
  analytes = c("acetic_acid", "propionic_acid", "butyric_acid"),
  stats = "anova",
  aov_formula = value ~ group
)
```

By default, the report workflow writes `scfa_concentration_normalized.csv` next
to the rendered HTML file.

## Main Functions

- `read_scfa_raw()` imports Chromeleon SCFA exports.
- `read_scfa_manifest()` imports sample metadata.
- `run_scfa_pipeline()` runs import, merge, normalization, summaries, models,
  and plotting.
- `render_scfa_report()` renders the Quarto report template.
- `default_scfa_analytes()` returns the default ordered SCFA analyte names.

## Development

Run the test suite with:

```r
devtools::test()
```

Regenerate `README.md` after editing `README.Rmd` with:

```r
devtools::build_readme()
```
